Idealized Wind Models of Cool Gas Outflows

### 2010-09-09 Winds/papers/Theory/First/Submit/ms.tex
blah
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5179 ac11df5e-a788-c23b-a8e1-c758c323924d

Starburst galaxies, whose luminosity is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  These
flows are generally expected (and sometimes observed) to have multiple
phases, for example a hot and diffuse phase traced by X-ray emission
together with a cool, denser phase traced by H$\alpha$ emission 
%% KHRR -- added refs
\citep[e.g.][]{ham90,martin99,shc+04,km10}. 
Several spectacular
examples in the local universe demonstrate that flows can extend to
up to $\sim 10$ kpc
%tens of kpc 
from the galaxy \citep{M87,lhw99,wsg08,vsr+03} carrying significant speed
to escape from the gravitational potential well of the galaxy's dark
matter halo \citep{wind_escape,sh09}.

@@

The observation of metal-line absorption is now a well-established
means of identifying outflows.  Furthermore, because the X-ray and
H$\alpha$ emission generated by winds is faint, absorption-line
analyses have traditionally been the only way to probe outflows in
distant galaxies. However, very little research
has been directed
toward comparing the observations against (even idealized) wind
models (although see \citealt{fmm+09}).  
Instead, researchers have gleaned what limited information
is afforded by direct analysis of the absorption lines.  The data do yield
robust measurements of the speed of the gas (is this really true?), 
yet they poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the flow from the galaxy.   In turn, constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of
galactic-scale winds, therefore, remain open matters of debate
\citep{debate,mqt05,sdr08,sh09,ssr02,od06,kkd+09}.

@@

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite{rubin+10c}
and Rubin et al.\ 2011 (in prep).
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attribute
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P
Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions \citep{prs+02,shapley03}.
We emphasize that this emission arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

@@

Although astronomers are rapidly producing a wealth of observational
datasets on galactic-scale winds, a key ingredient to a proper
analysis is absent.  Just as comparisons between observed supernova 
lightcurves and spectra and radiative transfer calculations of
of SNe models have provided crucial insight into the physics driving,
e.g., standard candle relationships for both Type Ia and IIP SNe 
\citep{kw07,kw09}, modeling of the observable signatures of galactic
outflows is necessary for understanding the physical
properties of the gas and ultimately the physics driving the flows. 
%[Draw analogy to exploring SN lightcurves and spectra]
In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows as observed in one
dimensional spectra and by IFUs.  Using Monte Carlo
radiative transfer techniques, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties, accounting for the effects of resonant scattering and
fluorescence for the first time.  
Although the winds are idealized, the results frequently
contradict our intuition and 
challenge the straightforward conversion of observables to (even crude) physical
constraints.  We caution that these findings have a direct bearing on
recent and upcoming surveys of galactic outflows, particularly those
which make use of \ion{Mg}{2} and \ion{Fe}{2} transitions to probe
outflow properties.
%[Add another line or two]

@@

@@

Focusing further on the \ion{Mg}{2} absorption, one notes that the profiles lie
well above zero intensity and have similar depth even though their $f\lambda$
values differ by a factor of two.  In standard absorption line
analysis, this is 
the tell-tale signature of a `cloud' that has a high optical depth (i.e.\
saturated) which only partially covers the emitting source
\citep[e.g.][]{rvs05a,hammann+10}.  Our fiducial wind model, however, 
{\it entirely covers the source}; the apparent partial covering must
be related to a different effect.
Figure~\ref{fig:noemiss} further emphasizes this point by comparing the 
absorption profiles from Figure~\ref{fig:fiducial_1d} against an
artificial model where no absorbed photons are 
re-emitted.   As expected from the
$\tau_{2796}$ profile (Figure~\ref{fig:fiducial_nvt}), this
`intrinsic' model
produces a strong \mgiid\ doublet that absorbs all photons at
$\delta v \approx -100 \mkms$, i.e.\ $I_{2796}^{\rm min} = \exp(-\tau^{\rm
  max}_{2796}) \approx 0$.
The true model, in contrast, has been `filled in' at $\delta v \approx -100
\mkms$ by photons scattered in the wind.  An
absorption-line analysis that ignores these effects
would (i) systematically underestimate the true optical
depth and (ii) falsely conclude that the wind partially covers the
source.  We will find that this
[behavior] is a generic result, even for wind models that include
dust and are not fully isotropic.

@@

[Add a paragraph on the bi-conical wind]

@@

Because of the high degree of photon trapping within the ISM
component, this model does suffer more from dust extinction than the
fiducial model.  We have studied the ISM+wind model including dust
with $\mtaud=1$ (normalized to include the ISM gas).  All of the 
emission lines are significantly reduced.  The \ion{Mg}{2} emission is
affect most because these lines are resonantly trapped.  The \feiis\
emission is also reduced relative to the dust-free model, but the
absolute flux still exceeds the fiducial model
(Table~\ref{tab:line_diag}).
[Should I show a figure?]

@@

Indeed, line-emission from low-ion transitions has been reported from 
star-forming galaxies that exhibit cool gas outflows.  
This includes emission related to \ion{Na}{1} \citep{phillips93,cth+10},
\ion{Mg}{2} emission \citep{wcp+09,rwk+10}, 
\ion{Si}{2}* emission from LBG galaxies \citep{shapley03},
and, most recently, significant \feiis\ emission \citep{rubin+10c}. 
A variety of origins have been proposed for this
line-emission including AGN activity, recombination in \ion{H}{2}
regions, and back-scattering off the galactic-scale wind.
[Are there aspects of those works that are now understood, e.g. the
difference in FeII 2586 and FeII 2600?]
Our results suggest that the majority of the observed line-emission is 
from scattered photons in the cool gas outflows of these star-forming
galaxies.  Indeed, our results indicate that line-emission may
be produced by all %of 
galaxies driving a wind.

On the other hand,
there are many examples of galaxies where blue-shifted absorption is
detected yet the authors report no significant line emission.  This
includes \ion{Na}{1} flows \cite{rvs05a,martin05,smn+09}, 
\ion{Mg}{2} and \ion{Fe}{2} absorption in
ULIRGS \citep{mb09}, and the extreme \ion{Mg}{2} outflows identified
by \cite{tmd07}.  Similarly, there have been no reports of
resonance line-emission from low-ion transitions in the
LBGs\footnote{\citet{psa+00}
  reported and analyzed \ion{C}{4} line-emission from the lensed LBG
  cb58, attributing this flux to stellar winds.  We caution, however,
  that the galactic-scale wind proposed to exist for cb58 should also
  have contributed to the \ion{C}{4} line-emission.}, and many
$z \lesssim 1$ galaxies show no detectable \ion{Mg}{2} or \ion{Fe}{2}*
emission despite significant blue-shifted absorption (Rubin et al., in prep). 
These non-detections appear to contradict a primary conclusion of 
this paper.   We are motivated, therefore, to
reassess several of the effects that can reduce the line-emission
and consider whether these could prevent its detection for many
star-forming galaxies.

Dust is frequently invoked to explain the suppression of line-emission
of resonantly trapped transitions (e.g.\ \lya).  Indeed, a photon
that is trapped for many scatterings within a dusty medium will 
preferentially be lost (???) relative to a non-resonant photon.  In
$\S$~\ref{sec:dust}, we examined the effects of dust extinction on 
\ion{Mg}{2} and \ion{Fe}{2} emission.  The general result
(Figure~\ref{fig:dust_tau}) is a modest
reduction in flux that scales as (1+\taud)$^{-1}$ instead of
exp(--\taud). Although the \ion{Mg}{2} photons are resonantly
trapped, they require only one to a few scatterings
to escape the wind model limiting the effects of
dust.  This reflects the moderate opacity of the \ion{Mg}{2}
doublet (e.g.\ relative to \lya) and also the velocity law of the
fiducial wind model.
As we have discussed, in scenarios where the \ion{Mg}{2} photons are
more effectively trapped dust further suppresses
the emission (e.g.\ the ISM+wind model from $\S$~\ref{sec:ISM}).
In contrast to \ion{Mg}{2}, 
the \ion{Fe}{2} resonance photons may be
converted to non-resonant photons that freely escape the wind.
Therefore, a wind model that trapped \ion{Mg}{2}
photons for many scatterings would not similarly trap 
\feiid\ photons.  The effects of dust are much 
reduced and, by inference, the same holds for any other set of
transitions that are coupled to a fine-structure level (e.g.\
\ion{Si}{2}/\ion{Si}{2}*).  In summary,
dust does reduce the line flux relative to the continuum, but 
it generally has only a modest effect on the predicted \ion{Mg}{2}
emission and always a minor effect on \feiis\ emission.

Another factor that may reduce line-emission is a
non-isotropic wind model. The flux is lower, for example, if one
eliminates the backside to the wind (e.g.\ the source itself could
shadow a significant fraction of the backside).  To zeroth order, the
line-emission scales 
roughly as $\Omega/4\pi$ where $\Omega$ is the angular extent of the
wind\footnote{There is also a first order dependence on the actual
  orientation of the wind, with maximal emission when the backside is
  fully viewed.}.  Because we require that the wind 
points toward us, it is difficult to reduce $\Omega$ much
below $2 \pi$ and, therefore, anisotropy likely reduces the
emission by at most a factor of order unity.  
Similarly, an anisotropic source may alter the predicted
line-emission.  For example, if the backside of the galaxy were
brighter/fainter then would one predict brighter/fainter emission
relative to the observed continuum.
To reduce the emission, the brightest regions
of the galaxy would need to be oriented toward Earth.  Although this
would not be a
generic orientation, most spectroscopic samples are magnitude-limited
and biased towards detecting galaxies when observed from their
brightest viewpoint.  In principle, this might imply a further
reduction of order unity.

Together, dust and anisotropic models may reduce the line-emission
of \ion{Mg}{2} and \ion{Fe}{2}
by one or two factors of order unity.  This would generally
not be sufficient, however, to reduce the emission below the typical detection
limits.

@@

We explore the effects of slit-loss as follows.
We model the slit as a perfect, infinitely-long rectangle centered on the
galaxy.  We then tabulate the equivalent width of the line-emission
through slits with a range of widths, parameterized by twice the radius
where the wind has a Sobolev optical depth of $\tau^S = 0.2$ (\rtt).
The line-emission will be weak beyond
this radius because the photons have only a low probability of
interacting with the wind.  For the fiducial wind model, 
the \mgiia\ transition has $\mrtt \approx 15$\,kpc 
(Figure~\ref{fig:fiducial_nvt}).  
The predicted equivalent width in emission \ewe\ relative to the absorption
equivalent width \ewabs\ is presented in
Figure~\ref{fig:obs_slit} for a series of transitions for the fiducial
wind model (black curves).
[Add LBG, radiative models]
The \ewe/\ewabs\ curves rise very steeply with increasing slit width and then
plateau when the slit width reaches $\approx \mrtt$.  
For this model, the emission is concentrated toward the source; this
derives from the density and velocity profiles but is also a simple
consequence of geometric projection.
The results are comparable for other wind models considered
in this paper (colored curves in Figure~\ref{fig:obs_slit}).
All of the curves rise steeply enough that 
the effects of slit-loss are minor (order unity)
unless the slit-width is very small.   
Nevertheless, the results do motivate
extended aperture observations, e.g.\ integral field unit (IFU)
instrumentation that would map the wind both spatially and spectrally.

Although several effects can reduce the line-emission
relative to the absorption of the outflow, our analysis
indicates that detectable line-emission should frequently occur. 
Furthermore, the line-emission could be suppressed so as not to exceed
the galaxy continuum yet still (partially) fill-in the absorption
profiles. Indeed, dust and anisotropic winds preferentially suppress
line-emission at $\mdv > 0 \mkms$
(Figures~\ref{fig:anisotropic},\ref{fig:dust}).
The emission would still 
modify the observed absorption profiles (e.g.\
Figure~\ref{fig:noemiss}) and may lead to erroneous conclusions
regarding characteristics of the flow.  {\it An analysis of cool gas outflows
that entirely ignores line-emission may incorrectly conclude that the gas is
partially covered, has a significantly lower peak optical depth,
and/or that a $\mdv \sim 0 \mkms$ component is
absent.}  
We now examine several quantitative effects of line-emission.

@@

These negative effects for \ion{Mg}{2} absorption are reduced
for the \ion{Fe}{2} absorption profiles, especially for \feiia.  This
is because a significant fraction (even a majority) of the
\ion{Fe}{2} emission is florescent \ion{Fe}{2}* emission at longer
wavelengths, which does not effect the absorption profiles. 
Therefore, the \ion{Fe}{2} absorption equivalent widths (\ewabs) more
closely follow the intrinsic value, one derives more accurate peak
optical depths, and the absorption kinematics more faithfully reflect
the motions of the flow.  Regarding the last point, one also
predicts a velocity offset between the \ion{Fe}{2} and \ion{Mg}{2}
centroids (Figure~\ref{fig:obs_vtau}).
%because scattered photons
%preferentially fill-in
%absorption at $v \sim 0 \mkms$ for the \ion{Mg}{2} transitions.  
This affects analysis of gas related to the ISM of the galaxy and
also material infalling at a modest speed.  One must take careful
consideration of line emission, therefore,  before drawing robust
conclusions about the presence/absence of gas with small velocity
offsets from systemic. Figure~\ref{fig:obs_vtau} also emphasizes that
the \ion{Mg}{2} profiles may misrepresent the kinematics of the bulk of
the gas.  Analysis of these lines, without consideration of
line-emission, would lead to incorrect conclusions on the energetics
and mass flux of the wind.

We emphasize that all of these effects are heightened by the
relatively low spectral resolution and S/N characteristic of the data
commonly acquired for $z>0$ star-forming galaxies.  Figure~\ref{fig:obs_lris}a
shows one realization of the \mgiid\ doublet for the ISM+wind model
convolved with the line-spread-function of the Keck/LRIS spectrometer
(a Gaussian with FWHM=250\kms) and an assumed signal-to-noise of
S/N=7 per (1 \AA) pixel.  Both the absorption
and emission are well detected, but it would be
difficult to resolve the issues discussed above (e.g.\ partial
covering, peak optical depth) with these data.  
One also notes several systematic effects of the lower spectral
resolution, e.g.\ reduced peak flux in the line-emission and a
systematic shift of \mgiia\ absorption to more negative velocity.
We have also modelled an observation of the ISM+wind model using
a stacked galaxy spectrum (Figure~\ref{fig:obs_lris}b).
Specifically, we
averaged 100 identical \ion{Mg}{2} profiles from the fiducial model
degraded to a S/N=2\,pix$^{-1}$ and shifted by a random velocity
offset with $\sigma = 100\mkms$.  
This treatment illustrates the effects of using stacked galaxy 
spectra to study outflows
\citep[e.g.][]{wcp+09,smn+09,rwk+10}.   The main difference from the
single galaxy observation shown in panel (a) is the smearing of
line-emission and absorption that reduces the height/depth of each.
The effects would be even more pronounced if one studied spectra with
a diversity of \ion{Mg}{2} profiles.    
Special care is required to interpret properties of the wind (and
ISM) from such spectral analysis.
[Emphasize that the ISM component has *no chance* of being revealed]
[{\bf Kate}: try stacking your set of observed LRIS data]

@@

The most obvious measures provided by line-emission analysis are
estimates
of the size and morphology of the outflow \citep{rubin+10c}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths.  The principle
challenge is to achieve sufficient sensitivity to detect low
surface brightness emission. As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds that at the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. [Give estimate]
[Add FeII* to the SB figure]
Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for this analysis because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  The \ion{Mg}{2} lines (especially \mgiia) has the
highest peak and integrated fluxes 
(Figure~\ref{fig:obs_sb}).

The kinematics of the line emission also offer insight into physical
characteristics of the wind.

[\feiis\ width!!]

>>>>

Starburst galaxies, whose luminosity is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  These
flows are generally expected (and sometimes observed) to have multiple
phases, for example a hot and diffuse phase traced by X-ray emission
together with a cool, denser phase traced by H$\alpha$ emission 
%% KHRR -- added refs
\citep[e.g.][]{ham90,martin99,shc+04,km10}. 
Several spectacular
examples in the local universe demonstrate that flows can extend to
up to $\sim 10$ kpc
%tens of kpc 
from the galaxy \citep{lhw99,wsg08,vsr+03} carrying significant speed
to escape from the gravitational potential well of the galaxy's dark
matter halo \citep[e.g.][]{sh09}.

@@

The observation of metal-line absorption is now a well-established
means of identifying outflows.  Furthermore, because the X-ray and
H$\alpha$ emission generated by winds is faint, absorption-line
analyses have traditionally been the only way to probe outflows in
distant galaxies. However, very little research
has been directed
toward comparing the observations against (even idealized) wind
models (although see \citealt{fmm+09}).  
Instead, researchers have gleaned what limited information
is afforded by direct analysis of the absorption lines.  The data
diagnose the speed of the gas relative to the galaxy, 
yet they poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the flow from the galaxy.   In turn, constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of
galactic-scale winds, therefore, remain open matters of debate
\citep{mqt05,sdr08,sh09,ssr02,od06,kkd+09}.

@@

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite[][see also
Rubin et al\ 2011, in prep.]{rubin+10c}.
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attribute
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P-Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line-emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions \citep{prs+02,shapley03}.
We emphasize that this emission arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

@@

Although astronomers are rapidly producing a wealth of observational
datasets on galactic-scale winds, a key ingredient to a proper
analysis is absent.  Just as comparisons between observed supernova 
lightcurves and spectra and radiative transfer calculations of
of SNe models have provided crucial insight into the physics driving,
e.g., standard candle relationships for both Type Ia and IIP SNe 
\citep{kw07,kw09}, modeling of the observable signatures of galactic
outflows is necessary for understanding the physical
properties of the gas and ultimately the physics driving the flows. 
%[Draw analogy to exploring SN lightcurves and spectra]
In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows as observed in one
dimensional spectra and by IFUs.  Using Monte Carlo
radiative transfer techniques, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties, accounting for the effects of resonant scattering and
fluorescence. 
Although the winds are idealized, the results frequently
contradict our intuition and 
challenge the straightforward conversion of observables to (even crude) physical
constraints.  We caution that these findings have a direct bearing on
recent and upcoming surveys of galactic outflows, particularly those
which make use of \ion{Mg}{2} and \ion{Fe}{2} transitions to probe
outflow properties.
%[Add another line or two]

@@

[JXP:  This needs to be re-run and revised following Kate's update.
Will do so in the next few days (hopefully tomorrow)]

@@

Focusing further on the \ion{Mg}{2} absorption, one notes that the profiles lie
well above zero intensity and have similar depth even though their $f\lambda$
values differ by a factor of two.  In standard absorption line
analysis, this is 
the tell-tale signature of a `cloud' that has a high optical depth (i.e.\
saturated) which only partially covers the emitting source
\citep[e.g.][]{rvs05a,hammann+10}.  Our fiducial wind model, however, 
{\it entirely covers the source}; the apparent partial covering must
be related to a different effect.
Figure~\ref{fig:noemiss} further emphasizes this point by comparing the 
absorption profiles from Figure~\ref{fig:fiducial_1d} against an
artificial model where no absorbed photons are 
re-emitted.   As expected from the
$\tau_{2796}$ profile (Figure~\ref{fig:fiducial_nvt}), this
`intrinsic' model
produces a strong \mgiid\ doublet that absorbs all photons at
$\delta v \approx -100 \mkms$, i.e.\ $I_{2796}^{\rm min} = \exp(-\tau^{\rm
  max}_{2796}) \approx 0$.
The true model, in contrast, has been `filled in' at $\delta v \approx -100
\mkms$ by photons scattered in the wind.  An
absorption-line analysis that ignores these effects
would (i) systematically underestimate the true optical
depth and/or (ii) falsely conclude that the wind partially covers the
source.  We will find that these are
generic results, even for wind models that include
dust and are not fully isotropic.

@@

[Add a paragraph on the bi-conical wind once we are confident in the
results.  This is may be a rather big effect.]

[Consider showing a 2D model for biconical]

@@

Because of the high degree of photon trapping within the ISM
component, this model does suffer more from dust extinction than the
fiducial model.  We have studied the ISM+wind model including dust
with $\mtaud=1$ (normalized to include the ISM gas).  All of the 
emission lines are significantly reduced.  The \ion{Mg}{2} emission is
affect most because these lines are resonantly trapped.  The \feiis\
emission is also reduced relative to the dust-free model, but the
absolute flux still exceeds the fiducial model
(Table~\ref{tab:line_diag}).  At large negative velocity offset from
systemic, the two profiles are nearly identical.

@@

Indeed, line-emission from low-ion transitions has been reported from 
star-forming galaxies that exhibit cool gas outflows.  
This includes emission related to \ion{Na}{1} \citep{phillips93,cth+10},
\ion{Mg}{2} emission \citep{wcp+09,rwk+10}, 
\ion{Si}{2}* emission from LBG galaxies \citep{shapley03},
and, most recently, significant \feiis\ emission \citep{rubin+10c}. 
A variety of origins have been proposed for this
line-emission including AGN activity, recombination in \ion{H}{2}
regions, and back-scattering off the galactic-scale wind.
Our results suggest that the majority of the observed line-emission is 
from scattered photons in the cool gas outflows of these star-forming
galaxies.  Indeed, line-emission should be generated by all 
galaxies driving a wind.

On the other hand,
there are many examples of galaxies where blue-shifted absorption is
detected yet the authors report no significant line emission.  This
includes \ion{Na}{1} flows \citep{rvs05a,martin05,smn+09}, 
\ion{Mg}{2} and \ion{Fe}{2} absorption in
ULIRGS \citep{mb09}, and the extreme \ion{Mg}{2} outflows identified
by \cite{tmd07}.  Similarly, there have been no reports of
resonance line-emission from low-ion transitions in the
LBGs\footnote{\citet{psa+00}
  reported and analyzed \ion{C}{4} line-emission from the lensed LBG
  cb58, attributing this flux to stellar winds.  We caution, however,
  that the galactic-scale wind proposed to exist for cb58 should also
  have contributed to the \ion{C}{4} line-emission.}, and many
$z \lesssim 1$ galaxies show no detectable \ion{Mg}{2} or \ion{Fe}{2}*
emission despite significant blue-shifted absorption (Rubin et al., in prep). 
These non-detections appear to contradict a primary conclusion of 
this paper.   We are motivated, therefore, to
reassess several of the effects that can reduce the line-emission
and consider whether these prevent its detection for many
star-forming galaxies.

Dust is frequently invoked to explain the suppression of line-emission
for resonantly trapped transitions (e.g.\ \lya).  Indeed, a photon
that is trapped for many scatterings within a dusty medium will 
be preferentially extincted relative to a non-resonant photon.  In
$\S$~\ref{sec:dust}, we examined the effects of dust extinction on 
\ion{Mg}{2} and \ion{Fe}{2} emission.  The general result
(Figure~\ref{fig:dust_tau}) is a modest
reduction in flux that scales as (1+\taud)$^{-1}$ instead of
exp(--\taud). Although the \ion{Mg}{2} photons are resonantly
trapped, they require only one to a few scatterings
to escape the wind which limits the effects of
dust.  This reflects the moderate opacity of the \ion{Mg}{2}
doublet (e.g.\ relative to \lya) and also the velocity law of the
fiducial wind model.
As we have discussed, in scenarios where the \ion{Mg}{2} photons are
more effectively trapped dust further suppresses
the emission (e.g.\ the ISM+wind model from $\S$~\ref{sec:ISM}).
In contrast to \ion{Mg}{2}, 
the \ion{Fe}{2} resonance photons may be
converted to non-resonant photons that freely escape the wind.
Therefore, a wind model that traps \ion{Mg}{2}
photons for many scatterings does not similarly trap 
\feiid\ photons.  The effects of dust are much 
reduced and, by inference, the same holds for any other set of
transitions that are coupled to a fine-structure level (e.g.\
\ion{Si}{2}/\ion{Si}{2}*).  In summary,
dust does reduce the line flux relative to the continuum, but 
it generally has only a modest effect on the predicted \ion{Mg}{2}
emission and always a minor effect on \feiis\ emission.

Another factor that may reduce line-emission is a
non-isotropic (e.g.\ biconical) wind model. The flux is lower, for example, if one
eliminates the backside to the wind (e.g.\ the source itself could
shadow a significant fraction of the backside).  
For our fiducial model, many of the photons scattered into our
sightline arise nearly normal to the sightline such that a biconicial
wind has signficantly lower line emission
($\S$~\ref{sec:anisotropic}). 
For the anisotropic winds explored in this paper,
the line-emission scales 
roughly as $\Omega/4\pi$ where $\Omega$ is the angular extent of the
wind\footnote{There is also a first order dependence on the actual
  orientation of the wind, with maximal emission when the backside is
  fully viewed.}.  Because we require that the wind 
points toward us, it is difficult to reduce $\Omega$ much
below $2 \pi$ and, therefore, anisotropy reduces the
emission by a factor of order unity.  
Similarly, an anisotropicly emitting source would modify the predicted
line-emission.  For example, if the backside of the galaxy were
brighter/fainter then would one predict brighter/fainter emission
relative to the observed continuum.
To reduce the emission, the brightest regions
of the galaxy would need to be oriented toward Earth.  Although this
would not be a
generic orientation, most spectroscopic samples are magnitude-limited
and biased towards detecting galaxies when observed from their
brightest viewpoint.  In principle, this might imply a further
reduction of order unity.

Together, dust and anisotropic models may reduce the line-emission
of \ion{Mg}{2} and \ion{Fe}{2}
by one to a few factors of order unity.  For some galaxies, these
effects may help to explain the absence of significant line-emission,
but we suspect these would generally
not be sufficient to reduce it below the detection limit.

@@

We explore the effects of slit-loss as follows.
We model the slit as a perfect, infinitely-long rectangle centered on the
galaxy.  We then tabulate the equivalent width of the line-emission
through slits with a range of widths, parameterized by \rtt: twice the radius
where the wind has a Sobolev optical depth of $\tau^S = 0.2$. 
The line-emission will be weak beyond
this radius because the photons have only a low probability of
being scattered.  For the fiducial wind model, 
the \mgiia\ transition has $\mrtt \approx 15$\,kpc 
(Figure~\ref{fig:fiducial_nvt}).  
The predicted equivalent width in emission \ewe\ relative to the absorption
equivalent width \ewabs\ is presented in
Figure~\ref{fig:obs_slit} for a series of transitions for the fiducial
wind model (black curves).
[Add radiative model]
The \ewe/\ewabs\ curves rise very steeply with increasing slit width and then
plateau when the slit width reaches $\approx \mrtt$.  
For the fiducial model, the emission is concentrated toward the source; this
derives from the density and velocity profiles but is also a simple
consequence of geometric projection.
The results are comparable for other wind models considered
in this paper. 
All of the curves rise steeply enough that 
the effects of slit-loss are minor (order unity)
unless the slit-width is very small.   
Nevertheless, the results do motivate
extended aperture observations, e.g.\ integral field unit (IFU)
instrumentation that would map the wind both spatially and spectrally.

Although several effects can reduce the line-emission
relative to the absorption of the outflow, our analysis
indicates that detectable line-emission should occur frequently. 
Furthermore, the line-emission could be suppressed so as not to exceed
the galaxy continuum yet still (partially) fill-in the absorption
profiles. Indeed, dust and anisotropic winds preferentially suppress
line-emission at $\mdv > 0 \mkms$
(Figures~\ref{fig:anisotropic},\ref{fig:dust}).
The emission would still 
modify the observed absorption profiles (e.g.\
Figure~\ref{fig:noemiss}) and may lead to erroneous conclusions
regarding characteristics of the flow.  {\it An analysis of cool gas outflows
that entirely ignores line-emission may incorrectly conclude that the gas is
partially covered, has a significantly lower peak optical depth,
and/or that a $\mdv \sim 0 \mkms$ component is
absent.}  
We now examine several quantitative effects of line-emission.

@@

These effects for \ion{Mg}{2} absorption are reduced
for the \ion{Fe}{2} absorption profiles, especially for \feiia.  This
is because a significant fraction (even a majority) of the
\ion{Fe}{2} emission is florescent \ion{Fe}{2}* emission at longer
wavelengths which does not affect the absorption profiles. 
Therefore, the \ion{Fe}{2} absorption equivalent widths (\ewabs) more
closely follow the intrinsic values, one derives more accurate peak
optical depths, and the absorption kinematics more faithfully reflect
the motions of the flow.  Regarding the last point, one also
predicts a velocity offset between the \ion{Fe}{2} and \ion{Mg}{2}
absorption-line centroids (Figure~\ref{fig:obs_vtau}).
%because scattered photons
%preferentially fill-in
%absorption at $v \sim 0 \mkms$ for the \ion{Mg}{2} transitions.  
This affects the analysis of gas related to the ISM of the galaxy and
also material infalling at a modest speed.  
Figure~\ref{fig:obs_vtau} also emphasizes that
the \ion{Mg}{2} profiles may misrepresent the kinematics of the bulk of
the gas.  Analysis of these lines, without consideration of
line-emission, would lead to incorrect conclusions on the energetics
and mass flux of the wind.

We emphasize that all of these effects are heightened by the
relatively low spectral resolution and S/N characteristic of the data
commonly acquired for $z>0$ star-forming galaxies.  Figure~\ref{fig:obs_lris}a
shows one realization of the \mgiid\ doublet for the ISM+wind model
convolved with the line-spread-function of the Keck/LRIS spectrometer
(a Gaussian with FWHM=250\kms) and an assumed signal-to-noise of
S/N=7 per (1 \AA) pixel.  Both the absorption
and emission are well detected, but it would be
difficult to resolve the issues discussed above (e.g.\ partial
covering, peak optical depth) with these data.  
One also notes several systematic effects of the lower spectral
resolution, e.g.\ reduced peak flux in the line-emission and a
systematic shift of \mgiia\ absorption to more negative velocity.
We have also modelled an observation of the ISM+wind model using
a stacked galaxy spectrum (Figure~\ref{fig:obs_lris}b).
Specifically, we
averaged 100 identical \ion{Mg}{2} profiles from the fiducial model
degraded to a S/N=2\,pix$^{-1}$ and shifted by a random velocity
offset with $\sigma = 100\mkms$.  
This treatment illustrates the effects of using stacked galaxy 
spectra to study outflows
\citep[e.g.][; S10]{wcp+09,rwk+10}.   The main difference from the
single galaxy observation shown in panel (a) is the smearing of
line-emission and absorption that reduces the height/depth of each.
The effects would be even more pronounced if one studied spectra with
a diversity of \ion{Mg}{2} profiles.    
Special care is required, therefore, to interpret properties of the wind (and
ISM) from such spectral analysis.

@@

The most obvious measures provided by line-emission analysis are
estimates
of the size and morphology of the outflow \citep{rubin+10c}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths.  The principle
challenge is to achieve sufficient sensitivity to detect low
surface brightness emission. As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds that at the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. {\bf [JXP: Give estimate]
[JXP: Add FeII* to the SB figure]}
Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for this analysis because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  The \ion{Mg}{2} lines (especially \mgiib) frequently has the
highest peak and integrated fluxes (Figure~\ref{fig:obs_sb}).  
On the other hand, the \ion{Mg}{2} transitions are more susceptible to
the effects of dust extinction and for some galaxies \feiis\ emission
will be dominant.  [Last comment]
